---
title: "Chi-Square Test for UX Researchers"
output: html_document
---

A chi-square test compares counts between groups. Here we use it to ask if the
new checkout design really changed how many users finish paying. Base R is
enough, no packages are needed.

## Make the data

We sample each user's outcome, giving the new design a better chance of
completing than the old design.

```{r data}
set.seed(11)
n <- 100

design  <- rep(c("Old", "New"), each = n)
old_out <- sample(c("Complete", "Abandon"), n, replace = TRUE, prob = c(0.40, 0.60))
new_out <- sample(c("Complete", "Abandon"), n, replace = TRUE, prob = c(0.62, 0.38))

data <- data.frame(design, outcome = c(old_out, new_out))
head(data)
```

## Build the table

`table()` counts how often each combination appears.

```{r table}
tbl <- table(data$design, data$outcome)
tbl
```

## Run the test

`chisq.test()` compares what we saw with what we would expect if design did
not matter.

```{r test}
test <- chisq.test(tbl)
test
test$p.value
```

## Check the assumptions

The test needs every expected count to be at least 5.

```{r expected}
test$expected
```

## Draw the completion rates

```{r plot}
completion <- prop.table(tbl, 1) * 100
barplot(completion[, "Complete"],
        names.arg = rownames(completion),
        ylab = "Completion (%)",
        main = "Completion rate by design")
```
