---
title: "Topic Modeling (LDA) on Support Tickets & User Reviews"
output: html_document
---

LDA reads many short customer texts and finds hidden groups of words that
appear together. Each group is a topic, so we can see the most common problems
in the reviews without reading every single one.

## Load packages

`quanteda` prepares the texts and `topicmodels` fits the LDA model.

```{r load}
library(quanteda)
library(topicmodels)
```

## Make some example reviews

We build each review from a random page and a random problem.

```{r data}
set.seed(42)

pages    <- c("camera", "login", "upload", "password", "profile")
problems <- c("crashes", "is very slow", "keeps failing", "freezes", "shows errors")

text <- paste(pages[sample(1:5, 12, replace = TRUE)],
              problems[sample(1:5, 12, replace = TRUE)])

data <- data.frame(id = 1:12, text = text)
data
```

## Split the texts into words

`tokens()` cuts a text into words and `tokens_remove()` deletes common words
like "the".

```{r tokens}
toks <- tokens(data$text, remove_punct = TRUE)
toks <- tokens_remove(toks, stopwords("en"))
toks
```

## Count words per review

A document-feature matrix (DFM) counts how often each word appears in each
review. LDA works on this matrix.

```{r dfm}
df_reviews <- dfm(toks)
df_reviews
```

## Fit the LDA model

We ask for 4 topics. `set.seed()` makes the result reproducible.

```{r lda}
set.seed(42)
model <- LDA(convert(df_reviews, to = "topicmodels"), k = 4)
model
```

## Look at the top words of each topic

`terms()` shows the 8 most important words per topic. Reading them names the
topic.

```{r top_words}
terms(model, 8)
```

## Count reviews per topic

`posterior()` gives the topic mix of each review. We label each review with
its strongest topic and count them.

```{r counts}
p <- posterior(model)
main_topic <- apply(p$topics, 1, which.max)
table(main_topic)
```
