---
title: "Quanteda Textual Analysis"
output: html_document
---

This notebook reads real texts from the corpus folder, cleans them, and asks
simple questions about the words: which words are most common and what the
whole corpus looks like.

## Load packages

```{r load}
library(quanteda)
library(readtext)
library(quanteda.textstats)
library(quanteda.textplots)
```

## Read the files in the corpus folder

`readtext()` imports every document into a data frame and shows the text.

```{r read}
files <- list.files("corpus", full.names = TRUE)
data <- readtext(files)
data$text
```

## Make a corpus

`corpus()` puts the texts into one quanteda object.

```{r corpus}
corp <- corpus(data, text_field = "text")
corp
```

## Split into words

`tokens()` cuts the texts into words and `tokens_remove()` deletes common
words like "the".

```{r tokens}
toks <- tokens(corp, remove_punct = TRUE)
toks <- tokens_remove(toks, stopwords("en"))
toks
```

## Count the words

`dfm()` makes a document-feature matrix that counts each word in each document.

```{r dfm}
dfmat <- dfm(toks)
dfmat
```

## Most frequent words

`textstat_frequency()` lists the most common words.

```{r freq}
textstat_frequency(dfmat, n = 10)
```

## Word cloud

`textplot_wordcloud()` draws the words with size by frequency.

```{r cloud}
textplot_wordcloud(dfmat, max_words = 40)
```
