---
title: "Confidence Intervals"
output: html_notebook
---

A notebook made by Paul Amat, [amat-design.com](amat-design.com)

## Context and objectives

> What is the confidence interval for... ?

## Data preparation and exploration

### Starting session

```{r}
set.seed(4218)
data <- data.frame(
  score  = rnorm(50, mean = 70, sd = 10),
  gender = sample(c("Man", "Woman"), 50, replace = TRUE)
)
```

## Confidence Intervals of Means

### Assuming Normality

```{r}
t.test(data$score)$conf.int
```

### Not assuming Normality

```{r}
wilcox.test(data$score, conf.int=TRUE)$conf.int
```

## Confidence Intervals of Proportion

```{r}
binom.test(length(data$gender[data$gender == "Man"]), length(data$gender))$conf.int
```

## Confidence Intervals of any computed value

### Boostraping

```{r}
boot <- numeric(1000) 
for(i in 1:1000) {
  boot[i]=mean(sample(data$score,10,replace=T)) 
}
```

### Histogram

```{r}
hist(boot)
```

### Displaying CI

```{r}
boot.sorted=sort(boot)
c(boot.sorted[25], boot.sorted[975])
```
