---
title: "Proportions differs in two groups"
output: html_notebook
---

A notebook made by Paul Amat, [amat-design.com](amat-design.com)

## Context and objectives

> Do two proportions measured in two groups differ from each other?

### Research hypothesis

### Statistical hypothesis

H0:

H1:

## Data preparation and exploration

### Starting session

```{r}
set.seed(6127)
data <- data.frame(
  gender = rep(c("Man", "Woman"), each = 60),
  mood   = c(sample(c("Happy", "Sad"), 60, replace = TRUE, prob = c(0.45, 0.55)),
             sample(c("Happy", "Sad"), 60, replace = TRUE, prob = c(0.65, 0.35)))
)
```

## Parametric test

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

#### Chaque effectif > 5

```{r}
table(data$gender, data$mood) > 5
```

### Test

```{r}
prop.test(table(data$gender, data$mood), correct=FALSE) # ok jamovi in contingency tables
```

### Effect size

## Non-parametric test

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

### Test

```{r}
fisher.test(table(data$gender, data$mood)) #ok jamovi
```

### Effect size

## Results and recommendations
