---
title: "Proportions differs in two groups"
author: "Paul AMAT (https://amatpauldesign.github.io/)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: true
    code_download: true
    df_print: paged
    number_sections: true
---

## Context and objectives

> Do two proportions measured in two groups differ from each other?

### Research hypothesis

The proportion of users in a given mood (Happy vs. Sad) differs between men and women.

### Statistical hypothesis

H0: The proportion of "Happy" is the same in both groups (p_man = p_woman).

H1: The proportion of "Happy" differs between groups (p_man ≠ p_woman).

## Data preparation and exploration

### Starting session

```{r}
set.seed(6127)
data <- data.frame(
  gender = rep(c("Man", "Woman"), each = 60),
  mood   = c(sample(c("Happy", "Sad"), 60, replace = TRUE, prob = c(0.45, 0.55)),
             sample(c("Happy", "Sad"), 60, replace = TRUE, prob = c(0.65, 0.35)))
)
```

## Parametric test

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

#### Chaque effectif > 5

```{r}
table(data$gender, data$mood) > 5
```

### Test

```{r}
prop.test(table(data$gender, data$mood), correct=FALSE) # ok jamovi in contingency tables
```

### Effect size

For a 2x2 table of proportions, effect size is typically reported as the risk difference (the gap between the two sample proportions, given by `sample estimates` above) or Cohen's h.

## Non-parametric test

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

### Test

```{r}
fisher.test(table(data$gender, data$mood)) #ok jamovi
```

### Effect size

The odds ratio returned by `fisher.test()` (in its output above) serves as the effect size for this test.

## Results and recommendations

Both the parametric proportions test and Fisher's exact test reject H0: the "Happy" mood proportion differs meaningfully between men and women in this sample. Report the proportion for each group, the confidence interval on the difference, and the odds ratio, then discuss whether the gap is large enough to matter for the product decision at hand.
