---
title: "Association between two independent nominal variables"
author: "Paul AMAT (https://amatpauldesign.github.io/)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: true
    code_download: true
    df_print: paged
    number_sections: true
---

## Context and objectives

> Is there an association between two nominal variables?

### Research hypothesis

Mood (Happy vs. Sad) is associated with gender.

### Statistical hypothesis

H0: Mood and gender are independent (no association).

H1: Mood and gender are associated.

## Data preparation and exploration

### Starting session

```{r}
set.seed(9043)
data <- data.frame(
  gender = rep(c("Man", "Woman"), each = 60),
  mood   = c(sample(c("Happy", "Sad"), 60, replace = TRUE, prob = c(0.40, 0.60)),
             sample(c("Happy", "Sad"), 60, replace = TRUE, prob = c(0.65, 0.35))),
  profil = sample(c("Athletic", "Dreamer", "Scientist"), 120, replace = TRUE)
)
```

### Observed counts

```{r}
t <- table(data$mood, data$gender)
t
```

### Visualize

```{r}
par(mfrow = c(1, 2))
barplot(t[,1], main = "Man")
barplot(t[,2], main = "Woman")
```

### Proportions

```{r}
round(prop.table(t, 1),2)
```

## Parametric test

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

### Test

```{r}
test <- chisq.test(data$mood, data$gender) #ok jamovi
test
```

### Post-test conditions

#### More than 5 observations in each combination of theoretical frequencies

```{r}
all(test$expected > 5)
```

### Effect size (nominal)

#### Standardized residuals

```{r}
test$stdres
```

Values greater than 2 indicate over-representation, values less than -2 indicate under-representation.

#### Odds ratio (2x2 only)

```{r}
n00 <- t[1,1]
n01 <- t[1,2]
n10 <- t[2,1]
n11 <- t[2,2]
OR <- (n00 * n11)/(n01 * n10)
OR # ok jamovi
```

Ratio of proportions. When there is no effect, the OR is equal to 1. In this case, the two proportions are equal: we have the same number in the numerator and in the denominator.

#### Relative risk (2x2 only)

```{r}
if(!require(DescTools)) install.packages("DescTools")
library(DescTools)

RelRisk(t) # ok jamovi
```

#### Cramer's V

```{r}
if(!require(rcompanion)) install.packages("rcompanion")
library(rcompanion)

(cramer <- cramerV(data$mood, data$gender)) #ok jamovi
```

Cramer's V for tables bigger than 2x2. Effect: .10 low, .30 moderate, .50 high.

### Effect size (ordinal)

#### Gamma

```{r}
data$gender <- factor(data$gender)
data$mood <- factor(data$mood, levels=c("Happy", "Sad"))

if(!require(DescTools)) install.packages("DescTools")
library(DescTools)

GoodmanKruskalGamma(table(data$mood, data$gender)) # ok jamovi
```

For ordinal variables, the Gamma if parametric. The effect is interpreted as a correlation coefficient, which ranges from -1 to +1. The closer the result is to 0, the less the variables are dependent on each other, and vice versa.

### Power analysis

#### Finding power

```{r}
if(!require(pwr)) install.packages("pwr")
library(pwr)
pwr.chisq.test(w = cramer, N = length(data$mood), df = test$parameter)
```

#### Finding N

```{r}
pwr.chisq.test(w = cramer, df = test$parameter, sig.level = 0.05, power = 0.8)
```

## Non-parametric test: Continuity correction

```{r}
test <- chisq.test(data$mood, data$gender, correct = T) #ok jamovi
test
```

Interpretation: continuity p-value from Yates or Fisher's exact test if theoretical numbers \<5. Continuity correction limits the bias brought by the fact that the sampling distribution is discrete but the chi-square distribution is continuous.

### Effect size (ordinal)

#### Kendall Tau B

```{r}
data$gender <- factor(data$gender)
data$mood <- factor(data$mood, levels=c("Happy", "Sad"))

if(!require(DescTools)) install.packages("DescTools")
library(DescTools)

KendallTauB(table(data$mood, data$gender)) # ok jamovi
```

For ordinal variables, Kendall's Tau-b if non-parametric. The effect is interpreted as a correlation coefficient, which ranges from -1 to +1. The closer the result is to 0, the less the variables are dependent on each other, and vice versa.

### Filter by adding a variable

```{r}
xtabs(~mood+gender+profil, data)
```


## Non-parametric test: Fisher's Exact Test

```{r}
fisher.test(data$mood, data$gender) #ok jamovi
```

## Results and recommendations

The chi-square test (with continuity correction) rejects independence between mood and gender in this sample, and the standardized residuals and Cramer's V above indicate the strength and direction of the association. Report which mood/gender combination is over- or under-represented, and confirm the finding holds under Fisher's exact test before drawing conclusions.
