---
title: "Means differ in two independant groups"
output:
  html_document:
    df_print: paged
---

A notebook made by Paul Amat, [amat-design.com](amat-design.com)

## Context and objectives

> Do these two groups differ?

### Research hypothesis

Means do not differ between groups/conditions.

### Statistical hypothesis

H0: mu1 = mu2

H1: mu1 ≠ mu2

## Data preparation and exploration

### Starting session

```{r}
set.seed(5286)
data <- data.frame(
  gender = rep(c("Man", "Woman"), each = 40),
  score  = c(rnorm(40, mean = 68, sd = 9),
             rnorm(40, mean = 74, sd = 9))
)
```

### Group descriptives

```{r}
group1 <- data$score[data$gender == "Man"]
group2 <- data$score[data$gender == "Woman"]

print("Man")
summary(group1)

print("Woman")
summary(group2)
```

### Mean difference

```{r}
mean(group1)-mean(group2) # ok jamovi
```

### Plot

```{r}
boxplot(score ~ gender, data)
```

## Parametric test

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

#### Normally distributed residuals

##### Residuals

```{r}
res <- residuals(lm(score ~ gender, data),type="response")
```

##### Plots

```{r}
par(mfrow=c(1,2))
hist(res, freq=FALSE)
curve(dnorm(x, mean=mean(res), sd=sd(res)), add=TRUE, col="red")
qqnorm(res)
qqline(res, col="red")
```

##### Shapiro-Wilk

```{r}
shapiro.test(res) # ok jamovi
```

#### Equal variances

##### F-test

```{r}
var.test(score ~ gender, data, conf.level=.95)
```

##### Bartlett

```{r}
bartlett.test(score ~ gender, data)
```

##### Levene

```{r}
library(car)
leveneTest(score ~ as.factor(gender), data, center = mean) # ok jamovi
```

In Levene test: if p-value is greater than 0.05, we do not reject H0 and we conclude that the variances of our two groups are homogeneous.

#### No outliers

```{r}
outliers <- function(x) {
   qinf <- quantile(x, 0.25) - (IQR(x)*1.5)
   qsup <- quantile(x, 0.75) + (IQR(x)*1.5)
   x[x < qinf | x > qsup]
}

outliers(group1)
outliers(group2)
```

### Test

```{r}
test <- t.test(score ~ gender, data, var.equal = T) # ok in jamovi
test
```

### Confidence intervals

```{r}
test$conf.int
```

This confidence interval encompasses the difference between the means of the groups. If the test is unilateral, one of the limits of the CI will be infinity.

### Effect size

```{r}
if(!require(effsize)) install.packages("effsize")
library(effsize)
(cohentest <- cohen.d(score ~ gender, data)) #ok jamovi
```

Cohen's d est faible si d = .20, modéré si d = .50, important si d = .80 . Le signe + ou - est donné par l'ordre des variables.

### Power analysis

#### Finding power

```{r}
if(!require(pwr)) install.packages("pwr")
library(pwr)

pwr.t.test(n = length(data$score)/2, 
           d = cohentest$estimate, 
           power = NULL, 
           type = "two.sample") # ok jamovi
```

#### Finding N

```{r}
pwr.t.test(n = NULL, 
           d = cohentest$estimate, 
           power = 0.9, 
           type = "two.sample") # ok jamovi
```

## Non-parametric test : non-equal variances

### Test

```{r}
t.test(score ~ gender, data, var.equal = F) # ok in jamovi
```

### Effect size

```{r}
if(!require(effsize)) install.packages("effsize")
library(effsize)
(cohentest <- cohen.d(score ~ gender, data, var.equal = F))
```

## Non-parametric test : non-normally distributed residuals

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

#### N in each group \> 20

```{r}
length(group1) > 20 & length(group2) > 20
```

### Test

```{r}
wilcox.test(score ~ gender, data, correct = F) # ok in jamovi
```

### Effect size

```{r}
if(!require(effectsize)) install.packages("effectsize")
library(effectsize)

rank_biserial(data$score, data$gender)
```

## Results and recommendations
