---
title: "SEQ & Task Time Joint Analysis"
output: html_document
---

For each task we know how hard users found it (SEQ, 1 = very hard, 7 = very
easy) and how long it took in seconds. Looking at both together shows which
tasks need the most work. Base R only, no packages needed.

## Make the data

We pick random values inside realistic ranges for SEQ and time for each task.

```{r data}
set.seed(6)

task <- c("Login", "Update profile", "Reset password", "Search results",
          "Checkout", "Upload photo", "Contact support", "Delete account")
seq  <- round(runif(8, 1, 7), 1)
time <- round(runif(8, 20, 500))

data <- data.frame(task, seq, time)
data
```

## Find the cutoffs

The median splits each measure into low and high.

```{r cutoffs}
median_seq  <- median(data$seq)
median_time <- median(data$time)

median_seq
median_time
```

## Label each task

Tasks that are slow and hard are frustrating. Tasks that are slow but feel
easy are invisible effort.

```{r classify}
data$group <- "Effortless"
data$group[data$time >  median_time & data$seq >= median_seq] <- "Invisible effort"
data$group[data$time >  median_time & data$seq <  median_seq] <- "Frustrating"
data$group[data$time <= median_time & data$seq <  median_seq] <- "Quick but annoying"

data
```

## Count the groups

```{r counts}
table(data$group)
```

## Draw time against SEQ

```{r plot}
plot(data$time, data$seq, col = as.factor(data$group), pch = 19,
     xlab = "Time (seconds)", ylab = "SEQ (1-7)")
abline(v = median_time, lty = 2)
abline(h = median_seq, lty = 2)
```
