---
title: "Means differ in more than two independant groups"
author: "Paul AMAT (https://amatpauldesign.github.io/)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: true
    code_download: true
    df_print: paged
    number_sections: true
---

## Context and objectives

> Is there at least one of the means or medians that differs from the others?

### Research hypothesis

Mean score differs across the three user profiles (Athletic, Dreamer, Scientist).

### Statistical hypothesis

H0: All group means are equal (mu_ath = mu_dre = mu_sci).

H1: At least one group mean differs from the others.

## Data preparation and exploration

### Starting session

```{r}
set.seed(3054)
data <- data.frame(
  profil = rep(c("Athletic", "Dreamer", "Scientist"), each = 30),
  score  = c(rnorm(30, mean = 70, sd = 8),
             rnorm(30, mean = 62, sd = 8),
             rnorm(30, mean = 75, sd = 8))
)
```

### Group descriptives

```{r}
levels(as.factor(data$profil))

score.ath <- data[which(data$profil == "Athletic"),]$score
score.dre <- data[which(data$profil == "Dreamer"),]$score
score.sci <- data[which(data$profil == "Scientist"),]$score

print("Athletic")
summary(score.ath)

print("Dreamer")
summary(score.dre)

print("Scientist")
summary(score.sci)
```

## Parametric test

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

#### Normally distributed residuals

##### Residuals

```{r}
res <- residuals(lm(score ~ profil, data),type="response")
```

##### Plots

```{r}
par(mfrow=c(1,2))
hist(res, freq=FALSE)
curve(dnorm(x, mean=mean(res), sd=sd(res)), add=TRUE, col = "red")
qqnorm(res)
qqline(res, col = "red")
```

##### Kolmogorov-Smirnov

```{r}
ks.test(res, "pnorm", mean(res), sd(res)) # ok jamovi
rm(res)
```

#### Equal variances

##### Levene

```{r}
car::leveneTest(score ~ as.factor(profil), data)
```

##### Bartlett

```{r}
bartlett.test(score ~ profil, data) # ok jamovi
```

##### Plots

```{r}
boxplot(score ~ profil, data)
```

#### No outliers

```{r}
outliers <- function(x) {
   qinf <- quantile(x, 0.25) - (IQR(x)*1.5)
   qsup <- quantile(x, 0.75) + (IQR(x)*1.5)
   x[x < qinf | x > qsup]
}

outliers(score.ath)
outliers(score.dre)
outliers(score.sci)
```

### Test

```{r}
model <- aov(score ~ profil, data)
summary(model) # ok jamovi
```

### Effect size

#### Eta-squared

```{r}
if(!require(lsr)) install.packages("lsr")
library(lsr)
etaSquared(model) # ok jamovi
```

The effect size for a one factor analysis of variance η2 (eta-squared) is the proportion of the variability of the outcome variable that can be explained by the predictor. If η2 = 0, there is no relationship between the two. If η2 = 1, the relationship is perfect.

The effect size for a multiple factors analysis η2p (partial eta-square, partial η2, or pη2) is the proportion of the variability of the outcome variable that can be explained by the predictor ignoring other effects in the model. If η2p = 0, there is no relationship between the two. If η2p = 1, the relationship is perfect. It is not possible to compare the partial values of η2p from one term to another.

#### Omega

```{r}

```

The square omega ω2 (omega-squared), less biased than η2, translates a weak effect if = 0.01, moderate if = 0.06, important if = 0.14.

### Tests post hoc

```{r}
pairwise.t.test(data$score, data$profil, p.adj = "bonferroni") # ok jamovi
```

For each post hoc t-test, the standard error rate I increases. For the total error rate after the tests to remain α = 0.05, the p-values must be adjusted.

If the condition of normality is respected, the adjusted p-value of HSD Tukey is to be read to see between which groups the difference is significant. The Bonferroni correction is to be used if the normality condition is not respected. This correction takes the initial threshold and divides it by the number of comparisons. The Holm correction is a more elaborate version of the Bonferroni correction.

#### Effect size for test post hoc

```{r}
score.ath <- which(data$profil == "Athletic")
score.dre <- which(data$profil == "Dreamer")
score.sci <- which(data$profil == "Scientist")

if(!require(effsize)) install.packages("effsize")
library(effsize)
(cohentest <- cohen.d(data$score[score.ath], data$score[score.dre]))
(cohentest <- cohen.d(data$score[score.ath], data$score[score.sci]))
(cohentest <- cohen.d(data$score[score.dre], data$score[score.sci])) # ok jamovi
```

## Non-parametric test: equal variances

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

### Test

```{r}
# fisher.test() is not appropriate here: score is continuous, so it would be
# treated as a ~90-level factor, producing a huge contingency table that
# makes the exact test computation hang. Fisher's exact test applies to
# categorical count data, not a continuous outcome.
```

### Post hoc test

```{r}
# Tukey
```

## Non-parametric test: unequal variances

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

### Test

```{r}
# welch
```

### Post hoc test

```{r}
# Games-Howell
```

## Non-parametric test

### Conditions

#### Independent observations

The independence of observations is guaranteed by the recruitment process.

### Test

```{r}
kruskal.test(score ~ profil, data) # ok jamovi
```

### Post hoc tests

```{r}
if(!require(PMCMRplus)) install.packages("PMCMRplus")
library("PMCMRplus")
dscfAllPairsTest(score ~ as.factor(profil), data = data) # ok jamovi
```

## Results and recommendations

The one-way ANOVA rejects H0: at least one profile's mean score differs from the others, with eta-squared above indicating the share of variance explained by profile. Use the Bonferroni-corrected pairwise t-tests (or the Dunn/DSCF post hoc test) to identify which specific pairs of profiles differ, and report those pairwise differences with their effect sizes.
