---
title: "Behavioral Persona Clustering with PAM"
author: "Paul AMAT (https://amatpauldesign.github.io/)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: true
    code_download: true
    df_print: paged
    number_sections: true
---

We have numbers about how users behave: how long their sessions last, how many
clicks they make, and how many features they use. PAM groups similar users
together so we can find the user types in the data instead of guessing.

## Load the package

`cluster` contains the `pam()` function.

```{r load}
library(cluster)
```

## Make some example users

We draw random values around three centers: power users, casual users and
feature-focused users.

```{r data}
set.seed(3)
n_each <- 4

duration <- c(rnorm(n_each, 30, 2), rnorm(n_each, 5, 2), rnorm(n_each, 16, 2))
clicks   <- round(c(rnorm(n_each, 60, 5), rnorm(n_each, 8, 3), rnorm(n_each, 22, 4)))
features <- c(rnorm(n_each, 7, 0.5), rnorm(n_each, 2, 0.3), rnorm(n_each, 4.5, 0.5))

data <- data.frame(duration, clicks, features)
data
```

## Put all variables on the same scale

`scale()` centers the columns so that `clicks`, which has big numbers, does
not dominate the distance.

```{r scale}
users_scaled <- scale(data)
users_scaled
```

## Group the users into 3 clusters

`pam()` chooses 3 real users as cluster centers and puts every user with the
nearest one.

```{r pam}
set.seed(1)
model <- pam(users_scaled, 3)
data$cluster <- model$clustering
data
```

## Pick the number of clusters

The average silhouette width measures how well users sit inside their cluster.
We try several values of k and keep the one with the highest score.

```{r silhouette}
set.seed(1)
sil_width <- sapply(2:6, function(k) pam(users_scaled, k)$silinfo$avg.width)
sil_width
```

## Describe each cluster

The average of each variable per cluster shows what kind of user it is.

```{r centers}
aggregate(data[, 1:3], by = list(cluster = data$cluster), mean)
```

## Draw the clusters

```{r plot}
plot(data$duration, data$clicks, col = data$cluster, pch = 19,
     xlab = "Duration", ylab = "Clicks")
```
