---
title: "Customer Feedback Sentiment Analysis with tidytext"
output: html_document
---

Sentiment lexicons give each word a score that shows how positive or negative
it is. Adding up the scores of an answer tells us if the feedback is happy or
angry, so we can compare releases without reading every comment.

## Load packages

`tidytext` splits texts into words and gives us the sentiment lexicons.

```{r load}
library(tidytext)
library(dplyr)
```

## Make some example feedback

We pick a random positive or negative comment for each customer and spread
them over three releases.

```{r data}
set.seed(7)

positive <- c("love the redesign, fast and easy",
              "great update, no crashes",
              "best app so far, a joy to use")
negative <- c("the checkout is confusing and slow",
              "app crashes every time, terrible",
              "i hate the new layout")

n <- 12
release <- rep(c("1.0", "1.1", "2.0"), each = 4)
mood    <- sample(c("positive", "negative"), n, replace = TRUE)
text    <- ifelse(mood == "positive",
                  sample(positive, n, replace = TRUE),
                  sample(negative, n, replace = TRUE))

data <- data.frame(id = 1:n, release = release, text = text)
data
```

## Split into words

`unnest_tokens()` puts one word per row and `anti_join()` removes stop words
(the, and, of...).

```{r words}
words <- unnest_tokens(data, word, text)
words <- anti_join(words, stop_words)
words
```

## Add an AFINN score to each word

The AFINN lexicon scores words from -5 (bad) to +5 (good).

```{r scores}
afinn <- get_sentiments("afinn")
scored <- inner_join(words, afinn, by = "word")
scored
```

## Average score per release

`aggregate()` computes the mean score for each release.

```{r by_release}
by_release <- aggregate(value ~ release, data = scored, mean)
by_release
```

## Count positive and negative words

The Bing lexicon simply labels words positive or negative.

```{r bing}
bing <- get_sentiments("bing")
pos_neg <- inner_join(words, bing, by = "word")
table(pos_neg$sentiment)
```
