---
title: "Customer Journey Sequence Analysis & Visualization"
author: "Paul AMAT (https://amatpauldesign.github.io/)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: true
    code_download: true
    df_print: paged
    number_sections: true
---

TraMineR draws the route each user takes through an app, step by step. Each
route is a sequence of pages, and the plots show the common paths and where
users stop.

## Load the package

```{r load}
library(TraMineR)
```

## Make some example journeys

Each user follows the stages in order, and each one stops at a random point.

```{r data}
set.seed(5)

stage <- c("S", "B", "P", "C", "E")
data  <- matrix(NA, nrow = 10, ncol = 5)

for (i in 1:10) {
  keep <- sample(2:5, 1)            # how far this user goes
  data[i, 1:keep] <- stage[1:keep]
}

data <- as.data.frame(data)
names(data) <- paste0("t", 1:5)
data
```

## Turn the table into a sequence object

`seqdef()` tells TraMineR the possible states and their names.

```{r seqdef}
alphabet <- c("S", "B", "P", "C", "E")
labels   <- c("Landing", "Browse", "Product", "Cart", "Order done")

journey <- seqdef(data, alphabet = alphabet, labels = labels)
journey
```

## Draw each user's journey

`seqiplot()` draws one coloured row per user from first to last step.

```{r seq_plot}
seqiplot(journey)
```

## Show the most common journeys

`seqfplot()` draws the paths that many users share.

```{r seqf_plot}
seqfplot(journey)
```

## Show the share of users in each state

`seqdplot()` shows, at each step, how many users are still in each state.

```{r seqd_plot}
seqdplot(journey)
```
