---
title: "Chi-Square Test for Checkout Completion"
author: "Paul AMAT (https://amatpauldesign.github.io/)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: true
    code_download: true
    df_print: paged
    number_sections: true
---

Does checkout completion depend on the device someone is using? We compare three device types (Desktop, iOS, Android) against task status (Pass / Fail) with a test of independence.

## Make the data

We sample each user's task outcome, giving Desktop the best odds of a
pass, Android the worst, and iOS in between.

```{r data}
set.seed(7361)
n <- 90

devices <- c("Desktop", "iOS", "Android")
status  <- c("Pass", "Fail")

desktop_out <- sample(status, n, replace = TRUE, prob = c(0.85, 0.15))
ios_out     <- sample(status, n, replace = TRUE, prob = c(0.70, 0.30))
android_out <- sample(status, n, replace = TRUE, prob = c(0.60, 0.40))

data <- data.frame(
  device = factor(rep(devices, each = n), levels = devices),
  status = factor(c(desktop_out, ios_out, android_out), levels = status)
)
head(data)
```

## Build the table

`table()` counts how often each combination appears. With three devices
and two outcomes, this is a 3x2 table.

```{r table}
tbl <- table(data$device, data$status)
tbl
```

## Run the test

`chisq.test()` compares what we saw with what we would expect if device
did not matter.

```{r test}
test <- chisq.test(tbl)
test
test$p.value
```

## Check the assumptions

The test needs every expected count to be at least 5.

```{r expected}
test$expected
```

## Visualize the counts with geom_count

```{r plot-counts}
library(ggplot2)

ggplot(data, aes(device, status)) +
  geom_count() +
  theme_minimal() +
  labs(x = "Device", y = "Task status", size = "Count",
       title = "Observed task outcomes by device")
```

## Visualize the residuals with geom_count

Residuals show which cells drove the test result: values far from
zero contributed the most to the chi-square statistic.

```{r plot-residuals}
resid_df <- as.data.frame(as.table(test$residuals))
names(resid_df) <- c("device", "status", "residual")

ggplot(resid_df, aes(device, status, color = residual)) +
  geom_text(aes(label = round(residual, 2)), size = 4) +
  scale_color_gradient2(low = "steelblue", mid = "gray70", high = "firebrick") +
  theme_minimal() +
  labs(x = "Device", y = "Task status", color = "Residual",
       title = "Pearson residuals by device")
```
