---
title: "UX Metrics Visualization"
author: "Paul AMAT (https://amatpauldesign.github.io/)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: true
    code_download: true
    df_print: paged
    number_sections: true
---

A set of standalone visualizations for common System Usability Scale (SUS) reporting needs: the overall score distribution, item-level breakdowns, subscale comparisons, demographic splits, and correlations. Each section simulates its own example data so the plots can be adapted independently.

```{r}
library(ggplot2)
```

## SUS Scores Distribution

```{r}
# Visual 1 DF: simulated overall SUS score per participant
set.seed(101)
df_vis1 <- data.frame(
  score = pmin(100, pmax(0, round(rnorm(150, mean = 72, sd = 15))))
)

ggplot(df_vis1, aes(x = score)) +
  geom_histogram(aes(fill = after_stat(x)), binwidth = 8) +
  coord_cartesian(xlim = c(0, 100)) +
  scale_fill_gradient(name = "Score", low = "red", high = "green", limits = c(0, 100)) +
  geom_vline(xintercept = mean(df_vis1$score), colour = "DodgerBlue") +
  geom_text(x = mean(df_vis1$score) + 1, y = 20, label = round(mean(df_vis1$score)), colour = "DodgerBlue", hjust = 0) +
  geom_vline(xintercept = 68, linetype = "dashed", colour = "#FF7F7F") +
  geom_text(x = 69, y = 22, label = "Market Benchmark (68)", colour = "#FF7F7F", hjust = 0) +
  labs(title = "Distribution of SUS Scores", subtitle = paste0("N = ", nrow(df_vis1)), x = "SUS Score", y = "Count")
```

## Items levels by subscales

```{r}
# Visual 2 DF: simulated mean score and 95% CI per SUS item
set.seed(202)
n_participants <- 100
item_mean <- round(runif(10, 60, 85))
item_sd <- runif(10, 10, 20)
item_se <- item_sd / sqrt(n_participants)

df_vis2 <- data.frame(
  Item = factor(paste("Item", 1:10), levels = paste("Item", 10:1)),
  Mean = item_mean,
  CI_low = item_mean - qt(0.975, n_participants - 1) * item_se,
  CI_high = item_mean + qt(0.975, n_participants - 1) * item_se
)
df_vis2$Subscale <- ifelse(df_vis2$Item %in% paste("Item", c(4, 10)), "Learnability", "Usability")

ggplot(df_vis2, aes(x = Mean, y = Item, fill = Subscale)) +
  geom_col(alpha = 0.7, width = 0.8) +
  geom_pointrange(aes(xmin = CI_low, xmax = CI_high), colour = "darkslategray") +
  geom_text(aes(x = 100, label = paste(round(Mean), "±", round(CI_high - Mean))), hjust = 1, colour = "darkslategray") +
  coord_cartesian(xlim = c(0, 100)) +
  theme_minimal() +
  labs(title = "SUS Item Mean Scores", x = "Mean Score (0–100 Scale)", y = "Item")
```

## 2D Subscales Matrix

```{r}
# Visual 3 DF: simulated usability vs. learnability subscale aggregates
set.seed(303)
df_vis3 <- data.frame(
  u_mean = rnorm(1, mean = 65, sd = 5),
  l_mean = rnorm(1, mean = 70, sd = 6),
  u_se = runif(1, 1, 3),
  l_se = runif(1, 1, 3)
)

ggplot() +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100)) +
  geom_point(aes(x = df_vis3$u_mean, y = df_vis3$l_mean), colour = "DodgerBlue", size = 3) +
  geom_rect(aes(
    xmin = df_vis3$u_mean - 1.96 * df_vis3$u_se, xmax = df_vis3$u_mean + 1.96 * df_vis3$u_se,
    ymin = df_vis3$l_mean - 1.96 * df_vis3$l_se, ymax = df_vis3$l_mean + 1.96 * df_vis3$l_se
  ), fill = "DodgerBlue", alpha = 0.2) +
  geom_point(aes(x = 68, y = 68), colour = "#FF7F7F", size = 3) +
  theme_light() +
  labs(title = "SUS Subscales: Usability vs. Learnability", x = "Usability", y = "Learnability")
```

## SUS scores by age groups

```{r}
# Visual 4 DF: simulated score paired with age group
set.seed(404)
df_vis4 <- data.frame(
  age_group = sample(c("18-24", "25-34", "35-44", "45-54", "55+"), 200, replace = TRUE),
  score = pmin(100, pmax(0, round(rnorm(200, mean = 70, sd = 15))))
)

ggplot(df_vis4, aes(x = score, y = age_group, color = score)) +
  geom_count() +
  scale_color_gradient(name = "Score", low = "red", high = "green", limits = c(0, 100)) +
  theme_minimal() +
  labs(title = "SUS Score x Age Group", x = "SUS Score", y = "Age Group")
```

## Chi-square residuals

```{r}
# Visual 5 DF: simulated contingency residuals between SUS grade category and age group
set.seed(505)
grade <- sample(
  c("F (<51)", "D/C (51-67)", "B (68-79)", "A (80+)"),
  200, replace = TRUE, prob = c(0.10, 0.25, 0.35, 0.30)
)
age_group <- sample(c("18-24", "25-34", "35-44", "45-54", "55+"), 200, replace = TRUE)

df_vis5 <- as.data.frame(chisq.test(grade, age_group)$residuals)
names(df_vis5) <- c("Grade", "AgeGroup", "Residuals")

ggplot(df_vis5, aes(x = Grade, y = AgeGroup, color = Residuals)) +
  scale_color_gradient2(low = "#FF4040", mid = "grey80", high = "#00B050", midpoint = 0) +
  geom_text(aes(label = round(Residuals, 2)), size = 4) +
  theme_minimal() +
  labs(title = "Chi-square Residuals: Grade Category x Age Group", x = "SUS Grade", y = "Age Group")
```

## SUS score by age correlation

```{r}
# Visual 6 DF: simulated numeric age vs overall SUS score
set.seed(606)
age <- round(runif(150, 18, 70))
score <- pmin(100, pmax(0, 50 + 0.3 * age + rnorm(150, mean = 0, sd = 15)))
df_vis6 <- data.frame(age = age, score = score)

r_val <- round(cor(df_vis6$age, df_vis6$score), 2)

ggplot(df_vis6, aes(x = age, y = score)) +
  geom_jitter(color = "darkslategray", alpha = 0.7) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, colour = "DodgerBlue") +
  annotate("text", x = mean(df_vis6$age), y = mean(df_vis6$score) + 5, label = paste("r =", r_val), colour = "DodgerBlue") +
  theme_minimal() +
  labs(title = "SUS Score x Continuous Age Correlation", x = "Age (Years)", y = "SUS Score")
```
